---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# rvtk

<!-- badges: start -->
[![R-CMD-check](https://github.com/astamm/rvtk/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/astamm/rvtk/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/astamm/rvtk/graph/badge.svg)](https://app.codecov.io/gh/astamm/rvtk)
<!-- badges: end -->

**rvtk** is an infrastructure package that makes the
[Visualization Toolkit (VTK)](https://vtk.org/) available to other R packages
that need to link against it. It provides three utility functions —
`CppFlags()`, `LdFlags()`, and `VtkVersion()` — that return the correct
compiler and linker flags for however VTK was found or installed on the
current machine.

## How VTK is located

On macOS and Linux the package runs a `configure` script at install time that
tries each of the following strategies in order, stopping as soon as one
succeeds:

1. The environment variable `VTK_DIR` (path to a VTK build or install tree).
2. [Homebrew](https://brew.sh/) (`brew --prefix vtk`).
3. `pkg-config` (`vtk-9.5`, `vtk-9.4`, …, `vtk-9.1`).
4. Common system prefixes (`/usr/local`, `/usr`, `/opt/local`).
5. Download pre-built static libraries from
   <https://github.com/astamm/rvtk/releases>.

On Windows, pre-built UCRT64 static libraries are always downloaded
automatically from the same URL.

Configuration results are stored in `inst/vtk.conf` and read at run time by
`CppFlags()`, `LdFlags()`, and `VtkVersion()`.

## Installation

```r
# install.packages("pak")
pak::pak("astamm/rvtk")
```

A system VTK installation (≥ 9.1.0) is not required: if none is found the
package downloads pre-built static libraries automatically.

## Usage for downstream package developers

Add **rvtk** to the `Imports` field of your `DESCRIPTION`:

```
Imports: rvtk
```

Then call `rvtk::CppFlags()` and `rvtk::LdFlags()` from your
`src/Makevars` (macOS / Linux):

```makefile
PKG_CPPFLAGS = $(shell "$(R_HOME)/bin$(R_ARCH_BIN)/Rscript" -e "rvtk::CppFlags()")
PKG_LIBS     = $(shell "$(R_HOME)/bin$(R_ARCH_BIN)/Rscript" -e "rvtk::LdFlags()")
```

And from `src/Makevars.win` (Windows):

```makefile
PKG_CPPFLAGS = $(shell "$(R_HOME)/bin$(R_ARCH_BIN)/Rscript.exe" -e "rvtk::CppFlags()")
PKG_LIBS     = $(shell "$(R_HOME)/bin$(R_ARCH_BIN)/Rscript.exe" -e "rvtk::LdFlags()")
```

You can verify the detected installation at any time:

```{r}
library(rvtk)
CppFlags()
LdFlags()
VtkVersion()
```
