---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# rvtk

<!-- badges: start -->
[![R-CMD-check](https://github.com/astamm/rvtk/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/astamm/rvtk/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/astamm/rvtk/graph/badge.svg)](https://app.codecov.io/gh/astamm/rvtk)
<!-- badges: end -->

**rvtk** is an infrastructure package that makes the
[Visualization Toolkit (VTK)](https://vtk.org/) available to other R packages
that need to link against it. It provides three utility functions —
`CppFlags()`, `LdFlags()`, and `VtkVersion()` — that return the correct
compiler and linker flags for however VTK was found or installed on the
current machine.

## How VTK is located

On macOS and Linux the package runs a `configure` script at install time that
tries each of the following strategies in order, stopping as soon as one
succeeds:

1. The environment variable `VTK_DIR` (path to a VTK build or install tree).
2. [Homebrew](https://brew.sh/) (`brew --prefix vtk`).
3. `pkg-config` (`vtk-9.5`, `vtk-9.4`, …, `vtk-9.1`).
4. Common system prefixes (`/usr/local`, `/usr`, `/opt/local`).
5. Download pre-built static libraries from
   <https://github.com/astamm/rvtk/releases>.

On Windows, pre-built UCRT64 static libraries are always downloaded
automatically from the same URL.

> **Windows limitation:** The Rtools45 UCRT64 environment does not provide
> `netcdf` or `libproj`. Consequently, the following VTK modules are
> **disabled** in the Windows pre-built libraries: `VTK_IONetCDF`,
> `VTK_IOHDF`, `VTK_GeovisCore`, `VTK_RenderingCore`. Downstream packages
> that require any of these modules cannot currently be built on Windows with
> the pre-built libraries supplied by **rvtk**.

Configuration results are stored in `inst/vtk.conf` and read at run time by
`CppFlags()`, `LdFlags()`, and `VtkVersion()`.

## Installation

```r
# install.packages("pak")
pak::pak("astamm/rvtk")
```

A system VTK installation (≥ 9.1.0) is not required: if none is found the
package downloads pre-built static libraries automatically.

## Usage for downstream package developers

Add **rvtk** to the `Imports` field of your `DESCRIPTION`:

```
Imports: rvtk
```

Because `$(shell ...)` is a GNU make extension that is **not** allowed in
`Makevars`, the correct approach is to query `rvtk::CppFlags()` and
`rvtk::LdFlags()` from a `configure` / `configure.win` script and write the
results into `src/Makevars` at install time.

### Step 1 — `src/Makevars.in` (template, committed to version control)

```makefile
PKG_CPPFLAGS = @VTK_CPPFLAGS@
PKG_LIBS     = @VTK_LIBS@
```

### Step 2 — `configure`(`.win`)

```sh
# configure (macOS/Linux)
#!/bin/sh
set -e
: "${R_HOME:=$(R RHOME)}"
VTK_CPPFLAGS="$("${R_HOME}/bin/Rscript" --vanilla -e "rvtk::CppFlags()")"
VTK_LIBS="$("${R_HOME}/bin/Rscript" --vanilla -e "rvtk::LdFlags()")"
sed -e "s|@VTK_CPPFLAGS@|${VTK_CPPFLAGS}|g" \
    -e "s|@VTK_LIBS@|${VTK_LIBS}|g" \
    src/Makevars.in > src/Makevars
```

```sh
# configure.win (Windows)
#!/bin/sh
./configure
```

Make them executable:

```sh
chmod +x configure configure.win
```

### Step 3 — `.gitignore` / `.Rbuildignore`

Add generated `Makevars` to `.gitignore` so it is not committed:

```
src/Makevars
```

### Step 4 - `cleanup`(`.win`)

Add a `cleanup` / `cleanup.win` script that removes the generated `Makevars`
after installation so it is not accidentally committed:

```sh
# cleanup (macOS/Linux)
#!/bin/sh
rm -f src/Makevars
```

```sh
# cleanup.win (Windows)
#!/bin/sh
./cleanup
```

Make them executable:

```sh
chmod +x cleanup cleanup.win
```

## Querying the detected installation

You can verify the detected installation at any time:

```{r}
library(rvtk)
CppFlags()
LdFlags()
VtkVersion()
```
